---
title: 'R Project Samuele Ceol'
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)

`%notin%` <- Negate(`%in%`)
```

## Notes on translations

For the purpose of this analysis, it has been decided to provide a translation for all Italian terms that we will encounter.
Since many of this terms have a non-univocal translation in English, it has been decided to provide a translation table that goes as follows:

ITALIAN TERM                TRANSLATION
Scrutinio                   Ballot
Lista (elettorale)          (Electoral) list
Elettore                    Constituent
Votante                     Voter
Camera (dei Deputati)       Chamber (of Deputies)   
Senate (della Repubblica)   Senate (of the Republic)

In the following sections, during the length of the whole project, only the English terms will be used. 
The necessary steps will be taken in order to translate the Italian terms present in the used datasets to their English counterpart.

## Additional notes

In the context of the used datasets, the ballot contains information, for each election year and for both houses (Chamber and Senate), regarding the number of people that CAN vote (constituents) and the number of people that HAVE voted (voters).

The (electoral) list provides the number of votes received by each party during a given election year.

Italy introduced the universal suffrage in 1945 by extending the right to vote to women. 
Since the earliest election year in our dataset is 1948, all votes come from both male and female population.

All Italian citizen above the age of 18 can vote for the Chamber of Deputies.
Only those who have reached the age of 25 are able to vote for the election of the Senate.

## Importing and cleaning political data

In order to reduce the number of .csv files associated with the project, some operation have been performed on the data sets outside of RStudio. 
The .csv files on the Government website are separated by House (Chamber/Senate), by type (ballot/list) and by election year. In order to reduce the total number of election related files in the project, a new "Year" column has been added to the original data sets. After that the division by election year has been removed by merging the separate files.
The "candidate" column was also removed when present (since it was only there in an handful of election years).


```{r}
# TODO - ballot_chamber_it, "Valle d'aosta" contains an NA value for "Schede.non.valide" in 2013, for now we are going to count that value as zero

# National data (no region granularity)
ballot_senate_it <- read.csv(
  "./source/elections/italy/ballots-Senate.csv", 
  header = TRUE, 
  sep = ";"
)
ballot_chamber_it <- read.csv(
  "./source/elections/italy/ballots-Chamber.csv", 
  header = TRUE, 
  sep = ";"
)
list_senate_it <- read.csv(
  "./source/elections/italy/lists-Senate.csv",
  header = TRUE, 
  sep = ";"
)
list_chamber_it <- read.csv(
  "./source/elections/italy/lists-Chamber.csv", 
  header = TRUE, 
  sep = ";"
)

# Data with region granularity (No votes coming from outside of Italy)
ballot_senate_reg <- read.csv(
  "./source/elections/regions/ballots-Senate-Reg.csv", 
  header = TRUE, 
  sep = ";"
)
ballot_chamber_reg <- read.csv(
  "./source/elections/regions/ballots-Chamber-Reg.csv",
  header = TRUE, 
  sep = ";"
)
list_senate_reg <- read.csv(
  "./source/elections/regions/lists-Senate-Reg.csv", 
  header = TRUE, 
  sep = ";"
)
list_chamber_reg <- read.csv(
  "./source/elections/regions/lists-Chamber-Reg.csv", 
  header = TRUE, 
  sep = ";"
)

# Adding id column
ballot_senate_it <- ballot_senate_it %>% add_column(House = "Senate")
ballot_chamber_it <- ballot_chamber_it %>% add_column(House = "Chamber")
list_senate_it <- list_senate_it %>% add_column(House = "Senate")
list_chamber_it <- list_chamber_it %>% add_column(House = "Chamber")

ballot_senate_reg <- ballot_senate_reg %>% add_column(House = "Senate")
ballot_chamber_reg <- ballot_chamber_reg %>% add_column(House = "Chamber")
list_senate_reg <- list_senate_reg %>% add_column(House = "Senate")
list_chamber_reg <- list_chamber_reg %>% add_column(House = "Chamber")

# Joining 
ballot_it <- full_join(
  ballot_senate_it,
  ballot_chamber_it
)

list_it <- full_join(
  list_senate_it,
  list_chamber_it
)

ballot_reg <- full_join(
  ballot_senate_reg,
  ballot_chamber_reg
)

list_reg <- full_join(
  list_senate_reg,
  list_chamber_reg
)

# Translating column names
colnames(ballot_it) <- c("Region", "Constituents", "Voters", "Empty.votes", "Null.votes", "Year", "House")
colnames(list_it) <- c("Region", "Party", "Votes", "Year", "House")
colnames(ballot_reg) <- c("Region", "Constituents", "Voters", "Empty.votes", "Null.votes", "Year", "House")
colnames(list_reg) <- c("Region", "Party", "Votes", "Year", "House")

# We can group all lines that have the same year, house (and party for lists)
# We don't care if votes come from within our outside the country (Region column)
# We can also drop columns indicating empty (Schede.bianche) or invalid (Schede.non.valide) votes
# A new column indicating the voter turnout is also added
ballot_it <- ballot_it %>%
  subset(select = -c(Region, Empty.votes, Null.votes)) %>%
  group_by(Year, House) %>%
  summarise_all(sum) %>%
  mutate(Voter.turnout = format(round((Voters / Constituents) * 100, 1), nsmall = 1))

list_it <- list_it %>%
  subset(select = -Region) %>%
  group_by(Year, House, Party) %>%
  summarise_all(sum)

# We want to "standardize" region names for both regional lists and ballots
# Trim trailing whitespaces
ballot_reg$Region <- trimws(
    ballot_reg$Region, 
    which = "r"
  ) 

# For certain years, region are divided into "zones" (e.g. CAMPANIA 1, CAMPANIA 2, ...). We should remove the trailing integers
ballot_reg$Region <- sub(
    "\\s\\d+$", 
    "", 
    ballot_reg$Region
  )

# We can now group by region, year and house standardize names for regions in which is not constant throughout the whole dataset
ballot_reg <- ballot_reg %>%
  subset(select = -c(Empty.votes, Null.votes)) %>%
  group_by(Region, Year, House) %>%
  summarise_all(sum) %>%
  mutate(Voter.turnout = format(round((Voters / Constituents) * 100, 1), nsmall = 1)) %>%
  mutate(
    Region = str_to_upper(Region), Region = case_when(
      Region == "ABRUZZI" ~ "ABRUZZO",
      Region == "TRENTINO-ALTO ADIGE/SUDTIROL" ~ "TRENTINO ALTO ADIGE",
      Region == "TRENTINO-ALTO ADIGE" ~ "TRENTINO ALTO ADIGE",
      Region == "FRIULI VENEZIA GIULIA" ~ "FRIULI-VENEZIA GIULIA",
      Region == "EMILIA-ROMAGNA" ~ "EMILIA ROMAGNA",
      TRUE ~ Region
    )
  )

list_reg$Region <- trimws(
    list_reg$Region, 
    which = "r"
  ) 
list_reg$Region <- sub(
    "\\s\\d+$", 
    "", 
    list_reg$Region
  )

list_reg <- list_reg %>%
  group_by(Region, Year, House, Party) %>%
  summarise_all(sum) %>%
  mutate(
    Region = str_to_upper(Region), Region = case_when(
      Region == "ABRUZZI" ~ "ABRUZZO",
      Region == "TRENTINO-ALTO ADIGE/SUDTIROL" ~ "TRENTINO ALTO ADIGE",
      Region == "TRENTINO-ALTO ADIGE" ~ "TRENTINO ALTO ADIGE",
      Region == "FRIULI VENEZIA GIULIA" ~ "FRIULI-VENEZIA GIULIA",
      Region == "EMILIA-ROMAGNA" ~ "EMILIA ROMAGNA",
      TRUE ~ Region
    )
  )
```
For both national and regional lists we want to only keep a set percentile of top voted parties for each year.

```{r}
list_it_top <- list_it %>% 
  group_by(Year, House) %>% 
  filter(quantile(Votes, 0.7)<Votes)
  
list_reg_top <- list_reg %>% 
  group_by(Region, Year, House) %>% 
  filter(quantile(Votes, 0.7)<Votes)

ballot_it
list_it_top
ballot_reg
list_reg_top
```

## Importing and cleaning labour market data

First year with employment rate data (total + zone): 1973
First year with region granularity: 1993

We will start by importing data related to the employment rate in Italy.

```{r}
empl_rate_raw <- read.csv(
  "./source/labour/employment-rate.csv", 
  header = TRUE, 
  sep = ","
)

head(empl_rate_raw)
```

In the raw dataset we have many columns that have id codes that are spelled out in other columns

ID COLUMN       FULL NAME COLUMN
ITTER107        Territorio
TIPO_DATO_FOL   Tipo.dato
SEXISTAT1       Sesso
ETA1            Classe.di.eta
TITOLO_STUDIO   Titolo.di.studio
TIME            Seleziona.periodo 

We can drop all columns containing ids.
Using the unique() we can also notice that the columns "Tipo.data", "Flag.Codes" and "Flags" do not provide any useful data.

```{r}
# Drop
empl_rate <- empl_rate_raw %>%
  subset(select = -c(ITTER107, TIPO_DATO_FOL, SEXISTAT1, ETA1, TITOLO_STUDIO, TIME, Tipo.dato, Flag.Codes, Flags))

# Translate
colnames(empl_rate) <- c("Region", "Sex", "Age", "Education.level", "Year", "Value")
```

At this point, the dataset still contains an higher level of granularity than needed. 

Since we are not looking for a division based on the level of education, we want to keep only the rows where the education level has value total (totale), then we can drop the "Education.level" column.

We can do the same thing for the age column. In this case we will keep the rows that have age range 15-64, which is the widest possible range for people in their working years.
We could keep the 15+ range instead, but this would significantly lower the employment values since it also considers the senile population.

Totals values can also be kept for the sex related column.

The year columns also contains quarterly values for each year. We can drop the quarterly rows and keep only the aggregate data for each year.

We also need to drop some values from the Region column. In this regard we want to keep only the national values and the regional values (hence dropping the ones related to macro-regions).

Finally, we will provide a translation for the content of the Sex column.

```{r}
macro_regions <- c("Nord", "Nord-ovest", "Nord-est", "Centro", "Mezzogiorno")

#All quarterly values for the year column contain letters. We filter them out by trying to convert to integer and then checking if the resul is Na
empl_rate <- empl_rate %>%
  filter(
      Education.level == "totale" &
      Sex == "totale" &
      Age == "15-64 anni" &
      !is.na(as.numeric(Year)) &
      Region %notin% macro_regions
  ) %>%
  subset(select = -c(Education.level, Sex, Age))
```


Now, we can split this large dataset into a national and a regional one.

```{r}
empl_rate_it <- empl_rate %>%
  filter(Region == 'Italia') %>%
  subset(select = -Region)

empl_rate_reg <- empl_rate %>%
  filter(Region != 'Italia')
```








